---
title: "R_package_project"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R_package_project}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ADVTTEST)
```
# Theory

The use of t.test theory implemented in this package. Specifically, the package 
will test the NULL hypothesis that the underlying means of two populations are
the same. That is you will be testing:

$H_0 : \mu_x = \mu_y$

or more usefully

$H_1 :  \mu_x - \mu_y = 0$

This involves using a built-in function called `t.test(x,y)` where x and y are 
samples from the respective populations. Some important experimental design 
structures must be taken into consideration before using the function. These
considerations will be expressed through options supplied to the t.test function.

### Considerations

If x and y are independent and the variances in the populations are equal 
then the test is performed using `var.equal = TRUE`

`t.test(x,y, var.equal = TRUE)`

If x and y are independent and the population variances are not equal 
then the test is done using 

`t.test(x,y, var.equal = FALSE)`

If the samples are "paired", that is each experimental unit is measured 
twice, the first goes into x and the last into y then the appropriate test is

`t.test(x,y,paired=TRUE)`

The NULL hypothesis is placed in the options as $\mu=0$ which is the 
default value.


##  The myttest function 

Myttest function will investigate the data for evidence for a difference
in the population variances if the data supports the conclusion that the 
underlying variances are likely different or if paired = TRUE then an 
appropriate t-test is made to produce a list containing:

- "Welch", "T-test", "Paired" depending on what test was conducted
- A "Y/N" conclusion as to whether we should reject the NULL of equality 
of means at the alpha level of significance
- The summary statistics made by the appropriate t.test in a data list

The function will takes two vectors, alpha and paired. Assume that the 
variances in the population are the same (this will mean a simple no 
conditional function)

The function must produce a named list containing:

* A data frame containing x and y
* alpha $\alpha$
* paired
* The confidence interval for $\mu_x - \mu_y$
* The pvalue
* the test type
* the "Y/N" conclusion

The list will be of class "Rttest"

## My test Function
myttest function will look for a difference in the population variances and see whether or not the data.
If the data supports the conclusion that the variances are different or if the variances are paired then the t-test will be ran and will return a list.

- The one following type of test could be ran: "Welch", " T-test", "Paired"
- A decision on whether or not to reject the null of equality of the mean at the alpha level of significance.
- A summary of the chosen ttest.

```{r, eval=FALSE}
myttest = function(x, y, alpha, paired){

  # create a data frame
  df = data.frame(x = x, y = y)


  # finding the variance of the df
  var_list = sapply(df,var)

  # finding the population variance
  sd1 = sd(x)
  sd2 = sd(y)

  if(paired){
    print("Paired"); test_type = "Paired"; v = FALSE
  }
  else if (sd1 - sd2 <= 6) {
    print("T-test"); test_type = "T-test"; v = TRUE
  }
  else{
    print("WELCH "); test_type = "WELCH "; v = FALSE
  }

  # make t-test assuming that the variances in the population are the same
  test = t.test(x = x, y = y, var.equal = v, paired = paired, conf.level = 1 - alpha)

  # find the confidence intervals and p-value
  conf_inter = test$conf.int

  p_value = test$p.value

  if(p_value < alpha) {
    print("Y")
  }
  else{
    print("N")
  }

  # make a named list of function
  function_list = list(data = df, alpha = alpha, paired = paired,
                  conf_inter = conf_inter, p_value = p_value,
                  test_type = test_type)

  # make list to be of class "Rttest"
  class(function_list) = "Rttest"

  return(function_list)
}
```

## Print the information from my ttest
Print function prints to the command line  A (1-$\alpha$)100% confidence 
interval for $\mu_x - \mu_y$ and the type of test

```{r, eval=FALSE}
print.Rttest = function(x, ...) {

  ci = x$conf_inter  # store the confidence interval
  test_type = x$test_type   # store the type of test

  # create a list of summary data
  print_list = list(ci = ci, test_type = test_type)

  return(print_list)
}
```

## Creates an informative plot from the Rttest package

The "Rttest" class that creates boxplots of the data when 
non - paired data is analyzed. It also will make a boxplot of differences in 
data when `paired = TRUE` The ci for the difference of means is plotted inside
the boxplot of differences.

```{r, eval=FALSE}
plot.Rttest = function(x, ...){

  if(x$paired){

    data_var = NULL
    value    = NULL

    # merge the differences in data of x object
    m_data = data.frame(value = x$data$x - x$data$y, data_var = rep(LETTERS[24]))

    # create the boxplot when paired is True
    ggplot(m_data, aes(x = data_var , y= value , fill=data_var )) +
      geom_boxplot() +
      stat_summary(fun=mean, geom="point", shape=23, size=4, fill="white") +
      scale_fill_discrete(name = "categories") +
      labs(x = "Differences", y = "samples differences",
           title ="Boxplot of the samples")
  }
  else{

    data_var = NULL
    value    = NULL

    # create separate datasets of x and y
    x_data = data.frame(value = x$data$x, data_var = rep(LETTERS[24]))
    y_data = data.frame(value = x$data$y, data_var = rep(LETTERS[25]))

    # merge the x and y data into new data set
    merge_data = merge(x = x_data, y = y_data, all = TRUE)

    # create the boxplot
    ggplot(merge_data, aes(x = data_var , y= value , fill=data_var )) +
      scale_fill_discrete(name = "categories") +
      geom_boxplot() +
      labs(x = "category", y = "samples",
           title ="Boxplot of the samples")
  }
}
```
### Set data and call the ttest package I made
```{r}
set.seed(32); x=rnorm(30,mean=10,sd=15)
set.seed(35); y=rnorm(30,mean=8,sd=15)
ans1=ADVTTEST::myttest(x,y,alpha=0.05,paired=FALSE)
print(ans1)
plot(ans1)

set.seed(32); x=rnorm(30,mean=10,sd=5)
set.seed(35); y=rnorm(30,mean=8,sd=15)
ans2=ADVTTEST::myttest(x,y,alpha=0.05,paired=FALSE)
print(ans2)
plot(ans2)

set.seed(32); x=rnorm(30,mean=10,sd=15)
set.seed(35); y = x+ rnorm(30, 5 ,4)
ans3=ADVTTEST::myttest(x,y,alpha=0.05,paired=TRUE)
print(ans3)
plot(ans3)
```

